Staged Abstract Interpreters

### 2019-04-06 paper/60_cases.tex
paper

An commonly used technique to improve the running time in control-flow analysis
is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
implemented in monadic abstract interpreter by swaping the @StateT@ (for store)
and the @SetT@ (for nondeterminism). In this section, we show this optimization
is orthogonal to staging and we just need to swap the same \textit{staged}
version of monads. After the adjustment, we have the following monad stack:

@@

Abstract garbage collection is a technique to reclaim unreachable addresses in
the store while doing the abstract interpretation
\cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
shows that, for big-step monadic abstract interpreter, we may add a @Reader@
monad to track a set of root addresses and compute the reachable addresses every
time when we obtain a value.

\begin{lstlisting}
  def ask_root: AnsM[Set[Addr]]
  def eval(e: Expr): AnsM
\end{lstlisting}

>>>>

A commonly used technique to improve the running time in control-flow analysis
is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
implemented in monadic abstract interpreter by swaping the @StateT@ (for store)
and the @SetT@ (for nondeterminism). In this section, we show this optimization
is orthogonal to staging and we just need to swap the same \textit{staged}
version of monads. After the adjustment, we have the following monad stack:

@@

Abstract garbage collection is a technique to reclaim unreachable addresses in
the store while doing the abstract interpretation
\cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
shows that, for big-step monadic abstract interpreter, we may add a @Reader@
monad to track a set of root addresses and compute the reachable addresses every
time when we obtain a value. Here we adopt it with staged monads:

\begin{lstlisting}
  type AnsM[T] = RepReaderT[RepReaderT[
                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Store, ?],
                 Set[Addr], ?], Env, T] //add a reader monad for Set[Addr]
\end{lstlisting}

The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We modify the @fix@
function with caching (from Section \ref{sai}) to @fix_gc@, but the caching part
is omitted. Several auxiliary functions are used: @ask_root@ retrieves the
current root addresses from the moand stack; @root@ compute the reachable
addresses from a value; and @gc@ cleans up the store according to the reachable
addresses and returns a new store. For the generic interpreter, we also need to
compute the root addresses for compound expressions, readers may refer to
\cite{DBLP:journals/pacmpl/DaraisLNH17}.

\begin{figure}[h!]
\begin{lstlisting}
  def fix_gc(ev: (Expr => Ans) => (Expr => Ans))(e: Expr): Ans = for {
    ψ <- ask_roots ... // omit code for retrieving ρ, σ, in and out
    val res: Rep[(Set[(Value,Store)], Cache)] = ... // omit code for checking caches
    _ <- put_out_cache(res._2)
    vs <- lift_nd(res._1)
    σ <- gc(ψ ⊔ root(vs._2)) // perform abstract garbage collection
    _ <- put_store(σ)
  } yield v
\end{lstlisting}
\vspace{-1em}
\caption{\texttt{fix\_gc} for abstract garbage collection} \label{fixgc}
\end{figure}

%def root(v: Rep[Value]): Set[Addr]  def ask_root: AnsM[Set[Addr]]  def gc(ra: Rep[Set[Addr]]): AnsM[Store]
